Incorporating Gyroscope Information in Gesture Recognition

### 2011-12-17 paper/paper.tex
improved results section

	\subsubsection{Clustering}
		For all training example across all gesture types, kmeans defines a clustering function $f: \mathbb{R}^n \rightarrow \mathbb{Z}$. 
	\subsubsection{Hidden Markov Model}
		For each set of training examples for each gesture type train a Hidden Markov Model.  Namely in this stage, for each gesture type $g$, we generate a function $h_g: (\mathbb{Z}^k) \rightarrow \mathbb{R}_{[0,1]}$  by the Hidden Markov Model process:
$HMM: (\mathbb{Z}^k)^t \rightarrow h_g$, where each training example consists of k instantaneous data points and there are t training examples for gesture type g.  
	\subsubsection{Classification Procedure}
		To classify an unknown gesture example, which exists in $(\mathbb{R}^n)^k$, we first transform it to it’s dicsretized version, namely $f($\mathbb{R}^n)^k$ and then for each gesture type $g$, apply $h_g((f(\mathbb{R}^n))^k)$ and classify the example according to the $g’$ were $h_g’$ returns maximum value.

\section{Results}\label{results}
\paragraph{Clustering approach}
Our initial approach of clustering on all training data together
worked reasonably well.  When running on all 5 sample gestures and
with the HMM enabled, we got an an accuracy of 89.6\%.

However, on the same data, our cluster-matching algorithm combined with
an HMM reaches 94.6\%.  We also observed similar improvements with examples,
as seen in the data.

\paragraph{HMM}
Although we found a few examples where the cluster-matching algorithm
benefits from using a Hidden Markov Model, in most real-world examples
there is little benefit.  When running on all of the training data,
we found that adding the HMM actually decreased accuracy slightly,
from 96.7\% to 94.6\%, which may be due to random fluctuations.

@@

\begin{center}
\begin{tabular}{ | l | l | l | l | l | }
  \hline
  Normalization & Features & Algorithm & Gestures Types & Accuracy \\ \hline
  Yes	&	Both  &	All & Cluster-matching with HMM (10) &	94.6\% \\ \hline
  Yes	&	Both	&	All & Original  &	89.6\% \\ \hline
  Yes	&	Accel &	All & Cluster-matching with HMM (10) &	92.0\% \\ \hline
  Yes	&	Accel &	All &	Original & 84.2\% \\ \hline
  Yes	&	Gyro  &	All & Cluster-matching with HMM (10) & 92.3\% \\ \hline
  Yes	&	Gyro  &	All & Original &  89.7\% \\ \hline
  Yes	&	Both	&	All &	Cluster-matching (10) &  96.7\% \\ \hline
  Yes	&	Both	&	All &	Cluster-matching (20) &  97.9\% \\ \hline
  Yes	&	Both  &	All &	Cluster-matching (40) & 	98.4\% \\ \hline
  No	&	Both	&	All &	Cluster-matching (40) & 	98.4\% \\ \hline
  Yes	&	Both	& Squares &	Cluster-matching (10) & 	100\% \\ \hline
  Yes	&	Both	& Circles and Reversed-circles  &	Cluster-matching (10) & 	Chance \\ \hline
  Yes	&	Both	& Circles and Reversed-circles  &	Cluster-matching with HMM (10) &	72.0\% \\ \hline
  Yes	&	Both  & Circles and Reversed-circles  &	Cluster-matching with HMM (20) & 88.8\% \\
  \hline
\end{tabular}
\end{center}

>>>>

	\subsubsection{Clustering}
		For all training example across all gesture types, kmeans defines a clustering function $f: \mathbb{R}^n \rightarrow \mathbb{Z}$. 
	\subsubsection{Hidden Markov Model}
		For each set of training examples for each gesture type train a Hidden Markov Model.  Namely in this stage, for each gesture type $g$, we generate a function $h_g: (\mathbb{Z}^k) \rightarrow \mathbb{R}_{[0,1]}$  by the Hidden Markov Model process:
$HMM: (\mathbb{Z}^k)^t \rightarrow h_g$, where each training example consists of k instantaneous data points and there are t training examples for gesture type g.  
	\subsubsection{Classification Procedure}
		To classify an unknown gesture example, which exists in $(\mathbb{R}^n)^k$, we first transform it to it’s dicsretized version, namely $f(\mathbb{R}^n)^k$ and then for each gesture type $g$, apply $h_g((f(\mathbb{R}^n))^k)$ and classify the example according to the $g’$ were $h_g’$ returns maximum value.

\section{Results}\label{results}

\begin{center}
\begin{tabular}{ | l | l | l | l | l | }
  \hline
  Normalization & Features & Algorithm & Gestures Types & Accuracy \\ \hline
  Yes	&	Both  &	All & Cluster-matching with HMM (10) &	94.6\% \\ \hline
  Yes	&	Both	&	All & Original  &	89.6\% \\ \hline
  No	&	Both	&	All & Original  &	88.4\% \\ \hline
  Yes	&	Accel &	All & Cluster-matching with HMM (10) &	92.0\% \\ \hline
  Yes	&	Accel &	All &	Original & 84.2\% \\ \hline
  Yes	&	Gyro  &	All & Cluster-matching with HMM (10) & 92.3\% \\ \hline
  Yes	&	Gyro  &	All & Original &  89.7\% \\ \hline
  Yes	&	Both	&	All &	Cluster-matching (10) &  96.7\% \\ \hline
  Yes	&	Both	&	All &	Cluster-matching (20) &  97.9\% \\ \hline
  Yes	&	Both  &	All &	Cluster-matching (40) & 	98.4\% \\ \hline
  No	&	Both	&	All &	Cluster-matching (40) & 	98.4\% \\ \hline
  Yes	&	Both	& Squares &	Cluster-matching (10) & 	100\% \\ \hline
  Yes	&	Both	& Circles and Reversed-circles  &	Cluster-matching (10) & 	Chance \\ \hline
  Yes	&	Both	& Circles and Reversed-circles  &	Cluster-matching with HMM (10) &	72.0\% \\ \hline
  Yes	&	Both  & Circles and Reversed-circles  &	Cluster-matching with HMM (20) & 88.8\% \\
  \hline
\end{tabular}
\end{center}

\subsection{Choice of features}
In general, the highest accuracy is achieved when using both accelerometer
and gyroscope data, followed by gyroscope data only, followed by accelerometer data only.

FINISH THIS

As we could not find any examples of using gyroscope data for gesture recognition
in the literature, this seems to be unique way of improving gesture recognition
accuracy.

\subsection{Clustering approach}
Our initial approach of clustering on all training data together
worked reasonably well.  When running on all 5 sample gestures and
with the HMM enabled, we got an an accuracy of 89.6\%.

However, on the same data, our cluster-matching algorithm combined with
an HMM reaches 94.6\% accuracy.  We also observed similar improvements with
other examples, as seen in the data.

\subsection{HMM}
Although we found a few examples where the cluster-matching algorithm
benefits from using a Hidden Markov Model, in most real-world examples
there is little benefit.  When running on all of the training data,
we found that adding the HMM actually decreased accuracy slightly,
from 96.7\% to 94.6\%, which may be due to random fluctuations.

@@

\subsection{Number of symbols}
Increasing the number of symbols in the HMM's alphabet improves accuracy
up to a point, but at the cost of a decrease in performance.

% For example, on the five original gestures, increasing the alphabet size
% from 10 to 40 improves the accuracy from 96.7\% to 98.4\%.  On the contrived
% circle data, the improvement was much larger, from 72.0\% to 88.8\%.

As shown in figure~\ref{fig:alphabetsize}, the optimum accuracy is achieved
at an alphabet size of around 25 symbols, with gradually decreasing accuracy
outside of that range.

\subsection{Normalization}
In general, our normalization makes only a small difference.  With the
original algorithm we saw an improvement of less than 1.5\% in accuracy,
and we saw no detectable difference using cluster-matching.
